---
title: "Long transient dynamics in the presence of noise - Figure 2"
author: "Jody R. Reimer"
date: "`r Sys.Date()`"
output: 
pdf_document: default
---

```{r library, message=F}
library(tidyverse)
library(nimble)
library(ecp)
library(viridis)
library(cowplot)
```
  
Define the model parameters. 

```{r}

n.its <- 5000 # number of iterations 
p <- list(r = .05, K = 2, Q = 5, H = .38, sigma = 0, a=0.023, N = 2.7e3, t.step = 1) # define parameter list

growth <- function(x, p) x * p$r * (1 - x / p$K) # growth terms in model
consumption <- function(x,p) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q) # consumption terms in model
```

Define the stochastic model in BUGS notation. 

```{r, message=FALSE}

may  <- nimble::nimbleCode({
  x[1] <- x0
  for(t in 1:((1/t.step)*N-1)){
    # calculate determinstic mean at next time step
    mu[t] <- x[t] + t.step*(x[t] * r * (1 - x[t] / K)  - a * x[t] ^ Q / (x[t] ^ Q + H ^ Q))
    # now add noise to that mean 
    y[t+1] ~ dnorm(mu[t], sd = sigma*mu[t]*sqrt(t.step)) 
    # note: since variance scales linearly with time, sd scales with the square root of time
    x[t+1] <- max(y[t+1],0) # truncate values so they don't go below 0
  }
})
model <- nimbleModel(may,constants = p, inits = list(x0 = 0.3)) # build model
cmodel <- model 
```

First consider sigma = 0 (i.e., the deterministic core, with no variance).

```{r, message=F}
simulate(cmodel)
df <- tibble(t = p$t.step*seq_along(cmodel$x), x.sigma0 = cmodel$x)
sigma0 = df
df %>% ggplot(aes(t, x.sigma0)) + geom_line() + theme_classic() + 
  theme(text = element_text(size = 20)) 
```

Next, set sigma = 0.005 and simulate n.its replicates, then take the mean. 

```{r, message=F}
p <- list(r = .05, K = 2, Q = 5, H = .38, sigma = 0.005, a=0.023, N = 2.7e3, t.step = 1)
model <- nimbleModel(may,constants = p, inits = list(x0 = 0.3))
cmodel <- model #compileNimble(model)
df <- map_dfr(1:n.its, 
  function(i){
    simulate(cmodel)
    tibble(t = p$t.step*seq_along(cmodel$x), x = cmodel$x, reps = i)
   })
df.reps <- df %>% spread(reps,x)
sigma0.005 = tibble(t = p$t.step*seq_along(cmodel$x),x.sigma0.005=rowMeans(df.reps[,-1]))
ggplot() + geom_line(data = df, aes(t,x,group=reps), alpha=.1) +
  geom_line(data = sigma0.005,aes(t,x.sigma0.005),color="yellow") + theme_classic() + 
  theme(text = element_text(size = 20)) 

```

Then do the same for sigma = 0.01; simulate n.ites replicates and take the mean 

```{r, message=F}
p <- list(r = .05, K = 2, Q = 5, H = .38, sigma = 0.01, a=0.023, N = 2.7e3, t.step = 1)
model <- nimbleModel(may,constants = p, inits = list(x0 = 0.3))
cmodel <- model #compileNimble(model)
df <- map_dfr(1:n.its, 
  function(i){
    simulate(cmodel)
    tibble(t = p$t.step*seq_along(cmodel$x), x = cmodel$x, reps = i)
   })
df.reps <- df %>% spread(reps,x)
sigma0.01 = tibble(t = p$t.step*seq_along(cmodel$x),x.sigma0.01=rowMeans(df.reps[,-1]))
ggplot() + geom_line(data = df, aes(t,x,group=reps), alpha=.1) +
  geom_line(data = sigma0.01,aes(t,x.sigma0.01),color="yellow",) + theme_classic() + 
  theme(text = element_text(size = 20)) 
```

Then for sigma = 0.015; simulate n.ites replicates and take the mean 

```{r, message=F}
p <- list(r = .05, K = 2, Q = 5, H = .38, sigma = 0.015, a=0.023, N = 2.7e3, t.step = 1)
model <- nimbleModel(may,constants = p, inits = list(x0 = 0.3))
cmodel <- model #compileNimble(model)
df <- map_dfr(1:n.its, 
  function(i){
    simulate(cmodel)
    tibble(t = p$t.step*seq_along(cmodel$x), x = cmodel$x, reps = i)
   })
df.reps <- df %>% spread(reps,x)
sigma0.015 = tibble(t = p$t.step*seq_along(cmodel$x),x.sigma0.015=rowMeans(df.reps[,-1]))
ggplot() + geom_line(data = df, aes(t,x,group=reps), alpha=.1) +
  geom_line(data = sigma0.015,aes(t,x.sigma0.015),color="yellow",) + theme_classic() + 
  theme(text = element_text(size = 20)) 
```

And finally for sigma = 0.02; simulate n.ites replicates and take the mean 

```{r, message=F}
p <- list(r = .05, K = 2, Q = 5, H = .38, sigma = 0.02, a=0.023, N = 2.7e3, t.step = 1)
model <- nimbleModel(may,constants = p, inits = list(x0 = 0.3))
cmodel <- model #compileNimble(model)
df <- map_dfr(1:n.its, 
  function(i){
    simulate(cmodel)
    tibble(t = p$t.step*seq_along(cmodel$x), x = cmodel$x, reps = i)
   })
df.reps <- df %>% spread(reps,x)
sigma0.02 = tibble(t = p$t.step*seq_along(cmodel$x),x.sigma0.02=rowMeans(df.reps[,-1]))
ggplot() + geom_line(data = df, aes(t,x,group=reps), alpha=.1) +
  geom_line(data = sigma0.02,aes(t,x.sigma0.02),color="yellow",) +
  theme_classic() + theme(text = element_text(size = 20)) 
```

Now plot all of the mean trajectories together.

```{r, message=F}
allsigmas = merge(sigma0, sigma0.005, by="t")
allsigmas = merge(allsigmas, sigma0.01, by="t")
allsigmas = merge(allsigmas, sigma0.015, by="t")
allsigmas = merge(allsigmas, sigma0.02, by="t")
colnames(allsigmas) = c("t","0","0.005","0.01","0.015","0.02")
allsigmasMelted <- reshape2::melt(allsigmas, id.var='t')
colnames(allsigmasMelted) = c("t","sigma","value")

ggplot(allsigmasMelted, aes(t, value)) + geom_line(aes(t, value,col=sigma), size=2) + scale_color_viridis(name = expression(sigma), discrete=TRUE, option = "D") + ylab("x") + theme_classic() + theme(text = element_text(size = 20)) +
  labs(subtitle = "(a)")  -> p1
p1 + theme(legend.position = c(0.8,0.3)) +
  theme(axis.title = element_text(size=25),
        axis.text = element_text(size=20),
        legend.text = element_text(size=20),
        legend.title = element_text(size=25),
        legend.title.align = 0.5,
        legend.text.align = 0)

p1 + theme(legend.position = c(0.8,0.3)) +
  theme(axis.title = element_text(size=25),
        axis.text = element_text(size=20),
        legend.text = element_text(size=20),
        legend.title = element_text(size=25),
        legend.title.align = 0.5,
        legend.text.align = 0)-> p1
```

Now make plot of the potential curve, with three different initial conditions position on the curve in the region of the ghost attractor. 

```{r, message=F}

p <- list(r = .05, K = 2, Q = 5, H = .38, sigma = 0, a=0.023, N = 2.7e3, t.step = 1)
epsilon = 0.07
theory <- tibble(x = seq(0,p$K, length.out = 100)) %>%
  mutate(g = growth(x, p), 
         c = consumption(x, p)) %>%
  mutate(potential = - cumsum(g - c)) %>%
  gather(curve, y, -x, -potential) 
col3 = viridis(3)
theory %>%
  ggplot(aes(x, potential)) + 
  geom_line(lwd = 2) +
  annotate("point", x = 0.5, y = -0.11, size=5, shape="circle", color=col3[1]) + 
  annotate("point", x = 0.5-epsilon, y = -0.107, size=5, shape="circle", color=col3[2]) +
  annotate("point", x = 0.5+epsilon, y = -0.1107, size=5, shape="circle", color=col3[3]) + 
  #annotate("text", x = 0.5, y = -0.02, label = "initial \nconditions", size=8) + 
  annotate("text", x = 0.5, y = -0.08, label = expression(x[0]),size=8) + 
  annotate("text", x = 0.3, y = -0.17, label = expression(x[0]-epsilon),size=8) + 
  annotate("segment", x = 0.3, xend = 0.42, y = -0.16, yend = -0.12, colour = col3[2], size=1.5, arrow=arrow())+
  annotate("text", x = 0.7, y = -0.17, label = expression(x[0]+epsilon),size=8) + 
  annotate("segment", x = 0.7, xend = 0.58, y = -0.16, yend = -0.125, colour = col3[3], size=1.5, arrow=arrow())+
  theme_classic() + 
  theme(text = element_text(size = 20)) +
  labs(subtitle = "(b)") -> p2
p2 + theme(axis.title = element_text(size=25),
     axis.text = element_text(size=20))

p2 + theme(axis.title = element_text(size=25),
     axis.text = element_text(size=20)) -> p2

```

Calculate deterministic trajectories of curves starting at each of those three initial conditions. 
```{r}
model <- nimbleModel(may,constants = p, inits = list(x0 = 0.5))
cmodel <- model #compileNimble(model)
simulate(cmodel)
df.1 <- tibble(t = p$t.step*seq_along(cmodel$x), x.1 = cmodel$x)

model <- nimbleModel(may,constants = p, inits = list(x0 = 0.5-epsilon))
cmodel <- model #compileNimble(model)
simulate(cmodel)
df.2 <- tibble(t = p$t.step*seq_along(cmodel$x), x.2 = cmodel$x)

model <- nimbleModel(may,constants = p, inits = list(x0 = 0.5+epsilon))
cmodel <- model #compileNimble(model)
simulate(cmodel)
df.3 <- tibble(t = p$t.step*seq_along(cmodel$x), x.3 = cmodel$x)

allinits = merge(df.1, df.2, by="t")
allinits = merge(allinits, df.3, by="t")
#colnames(allinits) = c("t", "x_0", "x_0+epsilon", "x_0-epsilon")
allinitsMelted <- reshape2::melt(allinits, id.var='t')
colnames(allinitsMelted) = c("t","initial.condition", "value")
```

Plot the three trajectories. 

```{r}
ggplot(allinitsMelted, aes(t, value, col=initial.condition)) + 
  geom_line(size=2) + 
  scale_color_manual(name = "initial \ncondition", values = col3, labels = c(expression(x[0]), expression(x[0]-epsilon), expression(x[0]+epsilon))) +  ylab("x") + 
  theme_classic()+ 
  theme(text = element_text(size = 20)) +
  labs(subtitle = "(c)") -> p3
p3 + theme(legend.position = c(0.8,0.3)) + 
  theme(axis.title = element_text(size=25),
        axis.text = element_text(size=20),
        legend.text = element_text(size=20),
        legend.title = element_text(size=25),
        legend.title.align = 0.5,
        legend.text.align = 0)

p3 + theme(legend.position = c(0.8,0.3)) + 
  theme(axis.title = element_text(size=25),
        axis.text = element_text(size=20),
        legend.text = element_text(size=20),
        legend.title = element_text(size=25),
        legend.title.align = 0.5,
        legend.text.align = 0) -> p3
```

Combine all three plots into one figure and save. 

```{r}

plot_grid(
  p1,p2,p3,
  ncol = 3)

# save as png, eps, or pdf:

ggsave(filename="3plots.png",width=24,height=(5/8)*8)

setEPS()
postscript("3plots.eps",width=24,height=(5/8)*8)
plot_grid(
  p1,p2,p3,
  ncol = 3)
dev.off()

ggsave(filename="3plots.pdf",width=24,height=5)
```

Alternatively, can put the three plots into one column when plotting:

```{r}
plot_grid(
  p1,p2,p3,
  ncol = 1)

ggsave(filename="3plots_tall.png",width=8,height=3*(5/8)*8)

setEPS()
postscript("3plots_tall.eps",width=8,height=3*(5/8)*8)
plot_grid(
  p1,p2,p3,
  ncol = 1)
dev.off()

ggsave(filename="3plots_tall.pdf",width=8,height=15)
```
